Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Patterns in Australian Spatial Statistics

### 2025-07-10 paper/kobakian-cook-anzjs.qmd
fixed references - all titlecase now

  # Downloaded data
d <- read_xlsx("data/experiment-export.xlsx", sheet=2) %>%
  filter(!is.na(contributor)) %>%
  mutate(contributor = factor(contributor))

n_all_contributors <- d %>% count(contributor, sort=TRUE) %>% 
  summarise(n_contributors = length(contributor))

@@

# Check data set 
# Need to clean multiple entries, 48, 24
# remove duplicated entries due to submit button
d <- d %>% group_by(group, contributor, image_name) %>%
  slice(1) %>% ungroup() %>% 
  arrange(group, contributor, plot_order)

# Remove contributors who did not provide answers to most questions
keep <- d %>% count(contributor, sort = TRUE) %>% filter(n > 10)
d <- d %>% 
  filter(contributor %in% keep$contributor) %>%
  filter(contributor != "1234567890") # Testing id

# Remove contributors who did not provide any choices
# Or an insufficient amount of responses
bad_contribs <- d %>% group_by(contributor) %>% 
  #summarise(sum0 = sum(choice)) %>% 
  summarise(n_zero = sum(if_else(choice == 0, 1, 0))) %>%
  #filter(sum0 < 13) %>% 
  filter(n_zero > 8) %>%
  pull(contributor)

# Person is in list above
same_answer <- d %>% group_by(contributor) %>% 
  summarise(v = sd(choice))

@@

d <- d %>% 
  filter(!(contributor %in% bad_contribs))

n_contributors <- d %>% count(contributor, sort=TRUE) %>% 
  summarise(n_contributors = length(contributor))

@@

d <- d %>% mutate(certainty = factor(as.character(certainty),
  levels = c("1", "2", "3", "4","5"), ordered=TRUE))
```

```{r reps}
replicate <- tibble(image_name = c("aus_cities_12_geo.png", "aus_cities_12_hex.png", 
                                   "aus_cities_3_geo.png", "aus_cities_3_hex.png",
                                   "aus_cities_4_geo.png", "aus_cities_4_hex.png",
                                   "aus_cities_9_geo.png", "aus_cities_9_hex.png",
                                   "aus_nwse_2_geo.png", "aus_nwse_2_hex.png",
                                   "aus_nwse_3_geo.png", "aus_nwse_3_hex.png",
                                   "aus_nwse_5_geo.png", "aus_nwse_5_hex.png",
                                   "aus_nwse_6_geo.png", "aus_nwse_6_hex.png",
                                   "aus_three_12_geo.png", "aus_three_12_hex.png",
                                   "aus_three_5_geo.png", "aus_three_5_hex.png",
                                   "aus_three_8_geo.png", "aus_three_8_hex.png",
                                   "aus_three_9_geo.png", "aus_three_9_hex.png"),
                    replicate = c(1, 1, 2, 2, 3, 3, 4, 4, 
                                  1, 1, 2, 2, 3, 3, 4, 4,
                                  1, 1, 2, 2, 3, 3, 4, 4))
# Add rep info to data
d <- d %>% left_join(., replicate, by = "image_name")
```

```{r pdetection_group}
# Tidy for analysis
d <- d %>% 
  separate(image_name, c("nothing", "trend", "location", "type", "extra"), remove = FALSE) %>%
  select(-nothing, -extra) %>%
  mutate(location = as.numeric(location), 
    # detect measures the accuracy of the choice
         detect = ifelse(location == choice, 1, 0)) %>% 
  mutate(trend = case_when(
    trend == "nwse" ~ "NW-SE",
    trend == "cities" ~ "All Cities",
    trend == "three" ~ "Three Cities")) %>% 
  mutate(trend = fct_relevel(trend,"All Cities","Three Cities", "NW-SE")) %>% 
  mutate(type = case_when(
    type == "hex" ~"Hex.",
    TRUE~"Choro.")) %>% 
    mutate(detect_f = factor(detect, levels = c(1, 0), labels = c("Detected? Yes", "Detected? No")))

plots <- d %>% group_by(group, trend, type, location) %>%
  # pdetect measures the aggregated accuracy of the choices
  summarise(pdetect = length(detect[detect == 1])/length(detect)) 
```

@@

sa2 <- sa22011 %>% 
  filter(!st_is_empty(geometry)) %>% 
  filter(!state_name_2011 == "Other Territories") %>% 
  filter(!sa2_name_2011 == "Lord Howe Island")

SIR <- read_csv("data/SIR Downloadable Data.csv") %>% 
  filter(SA2_name %in% sa2$sa2_name_2011) %>% 
  dplyr::select(Cancer_name, SA2_name, Sex_name, p50) %>% 
  filter(Cancer_name == "Thyroid", Sex_name == "Females")
ERP <- read_csv("data/ERP.csv") %>%
  filter(REGIONTYPE == "SA2", Time == 2011, Region %in% SIR$SA2_name) %>% 
  dplyr::select(Region, Value)
# Alternative maps
# Join with sa2 sf object
sa2thyroid_ERP <- SIR %>% 
  left_join(sa2, ., by = c("sa2_name_2011" = "SA2_name")) %>%
  left_join(., ERP %>% 
              dplyr::select(Region, 
              Population = Value), by = c("sa2_name_2011"= "Region")) %>% 
  filter(!st_is_empty(geometry))
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  #filter(!is.na(Population)) %>% 
  filter(!sa2_name_2011 == "Lord Howe Island") %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf()

@@

## Check the projection uses long and lat
## Create centroids set
sa2centroids <- sa2 %>% 
  create_centroids(., "sa2_name_2011")
## Create hexagon grid
# For speed, consider if the buffer distance beyond the centroids is necessary
grid <- create_grid(centroids = sa2centroids,
                    hex_size = 0.15,
                    buffer_dist = 4, verbose = TRUE)
## Allocate polygon centroids to hexagon grid points
# This is a time consuming process for a large data set
aus_hexmap <- allocate(
  centroids = sa2centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.15,
  hex_filter = 8,
  focal_points = capital_cities,
  width = 25,
  verbose = TRUE)
save(aus_hexmap, file = "data/aus_hexmap.rda") 
}

@@

## Prepare to plot
fort_hex <- fortify_hexagon(data = aus_hexmap,
                            sf_id = "sa2_name_2011",
                            hex_size = 0.3) %>% 
            left_join(sa2thyroid_ERP %>%
            select(sa2_name_2011, SIR, p50))
fort_aus <- sa2thyroid_ERP %>%
  fortify_sfc()
## Make a plot
aus_hexmap_plot <- ggplot() +
  geom_polygon(aes(x = long,  y = lat,  
        group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey", 
               linewidth = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, 
                   fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal() +
  ggtitle("b. hexagon tile map")
```

@@

ggplot(expt_design, ) +
  geom_tile(aes(x=group, y=1, fill=type)) +
  geom_text(aes(x=group, y=1, label = paste("loc:", location))) + 
  facet_grid(replicate~trend, axes = "all", labeller = labeller(replicate = label_both, trend = label_both)) +
  ylab("") +
  scale_fill_manual("", values = type_colors) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

@@

var.sim1 <- predict(var.g.dummy, newdata = sa3_centroids, nsim = 12) %>% 
  left_join(sa3_centroids, ., by=c("longitude", "latitude"))

sa3_sims1_1 <- as_tibble(var.sim1) %>% 
  mutate_at(sims, ~map_dbl(1:nrow(sa3), spatial_smoother, 
                           values_vector = ., area_weight = 0.5, neighbours_list = sa3_neighbours))

@@

```{r}
#| label: fig-detect-compare
#| fig-cap: "The detection rates achieved by participants are contrasted when viewing the four replicates of the three trend models. Each point shows the probability of detection for the lineup display, the facets separate the trend models hidden in the lineup. The points for the same data set shown in a choroleth or hexagon tile map display are linked to show the difference in the detection rate."
#| fig-height: 4
#| fig-width: 6
#| out-width: 100%
# Detectability rate for each lineup (image)
d_smry <- d %>% group_by(trend, type, replicate) %>%
  # pdetect measures the aggregated accuracy of the choices
  summarise(pdetect = length(detect[detect == 1])/length(detect)) %>%
  ungroup()

# Numerical summary
diffs <- d_smry %>% spread(type, pdetect) %>%
  mutate(dif = `Hex.` - `Choro.`)

@@

```{r}
#| eval: false
# Check detection rate across groups
d %>% group_by(group) %>%
    # pdetect measures the aggregated accuracy of the choices
    summarise(pdetect = length(detect[detect == 1])/length(detect)) %>%
    ungroup()
```

@@

d %>% group_by(type, trend) %>%
  summarise(m = as.character(round(mean(detect), 2)),
      std.dev = as.character(round(sd(detect), 2))) %>% 
  mutate(std.dev = ifelse(std.dev == 0.5, "(0.50)", paste0("(", std.dev, ")"))) %>%   mutate(m = ifelse(m == 0.4, "0.40", m)) %>% 
  gather(stat, value, m, std.dev) %>% 
  pivot_wider(names_from = "trend", values_from = "value") %>% 
  arrange(type) %>% ungroup() %>% 
  mutate(Type = types) %>% select(Type, `NW-SE`, `Three Cities`, `All Cities`) %>%
  knitr::kable(., format = "latex", align = "lccc", booktabs = TRUE, 
    linesep = c("", "\\addlinespace"))
```

@@

detection_rates <- tidy(detect_fit) %>%
  mutate(detection_rates = round(exp(estimate)/(1+exp(estimate)),2)) %>%
  select(term, estimate, detection_rates) %>% pull(detection_rates)

@@

tidy(detect_fit) %>%
  mutate_at(.vars = c("estimate", "std.error"), round, 2) %>% 
  mutate(p.value = round(p.value, digits=2)) %>% 
  rowwise() %>%
  mutate(sig = case_when(
  p.value <= 0.001 ~ "$^{***}$",
  p.value <= 0.01 ~  "$^{**}$",
  p.value <= 0.05 ~  "$^{*}$",
  p.value <= 0.01 ~  "$^{.}$",
  TRUE ~ "$^{ }$")) %>%
  ungroup() %>% 
  filter(!is.na(std.error)) %>%
  mutate(term = glmer_terms) %>% 
  select(Term = term, 
    Est. = estimate,
    `Std. Err.` = std.error, 
    `P-val.` = p.value, 
    Sig. = sig) %>% 
  knitr::kable(format = "latex", escape = FALSE, align= "rrlrr", 
    booktabs = T, linesep = c("", "\\addlinespace", "", "\\addlinespace", ""))
```

@@

```{r}
#| label: fig-beeswarm
#| fig-cap: "The distribution of the time taken (seconds) to submit a response for each combination of trend, whether the data plot was detected, and type of display, shown using horizontally jittered dotplots. The coloured point indicates average time taken for each map type. Although some participants take just a few seconds per evaluation, and some take as much as 60 seconds, but there is very little difference in time taken between map types."
#| fig-height: 4
#| fig-width: 6
#| out-width: 100%
s <- d %>% group_by(type, trend, detect_f) %>%
  summarise(m=median(time_taken), 
            q1=quantile(time_taken, 0.25), 
            q3=quantile(time_taken, 0.75)) 
ggplot() + 
  geom_quasirandom(data=d, aes(x=type, y=time_taken), alpha=0.9) + 
  #geom_hline(data=s, aes(yintercept=m, color=type)) +
  geom_point(data=s, aes(x=type, y=m, color=type), size=3, alpha=0.7) +
  #geom_errorbar(data=s, aes(x=type, ymin=q1, ymax=q3, color=type), width=0.3, size=2) +
  scale_color_manual("", values = type_colors) +
  facet_grid(detect_f~trend) + 
  ylab("Time taken (seconds)") + xlab("") +
  theme(legend.position="none")
```

```{r}
#| eval: false
# Check groups and demographics
d %>% group_by(group) %>%   
  summarise(m=mean(time_taken), 
            s=sd(time_taken))
d %>% group_by(gender) %>%   
  summarise(m=mean(time_taken), 
            s=sd(time_taken))
d %>% group_by(age) %>%   
  summarise(m=mean(time_taken), 
            s=sd(time_taken))
d %>% group_by(education) %>%   
  summarise(m=mean(time_taken), 
            s=sd(time_taken))

@@

```{r}
# Check time difference between default choice
d_zeros <- d %>% mutate(zeros = if_else(choice == 0, 0, 1)) %>%
  group_by(zeros) %>%   
  summarise(m=median(time_taken), 
            mn=min(time_taken),
            mx=max(time_taken),
            q1=quantile(time_taken, 0.25), 
            q3=quantile(time_taken, 0.75),
            n=n())

d_zeros_detect <- d %>% mutate(zeros = if_else(choice == 0, 0, 1)) %>%
  group_by(zeros, detect_f) %>%   
  summarise(m=median(time_taken), 
            mn=min(time_taken),
            mx=max(time_taken),
            q1=quantile(time_taken, 0.25), 
            q3=quantile(time_taken, 0.75),
            n=n())
# Summarise across subjects
contributor_smry <- d %>% group_by(contributor) %>%   
  summarise(m_d=mean(detect), 
            s_d=sd(detect),
            m_t=mean(time_taken), 
            s_t=sd(time_taken)) 
```

@@

```{r}
#| label: fig-certainty 
#| fig-cap: "The distribution of certainty chosen by participants when viewing hexagon tile map or choropleth displays, shown as a mosaic plot, faceted by the trend model and whether the plot was detected or not. At most trend models and whether the plot was detected or not, participants were more likely to choose a high certainty when evaluating a choropleth map."
#| fig-height: 5
#| fig-width: 10
#| out-width: 100%
d <- d %>% 
  mutate(certainty = as_factor(certainty)) %>% 
  mutate(replicate_f = as_factor(replicate))

#d %>% 
#  mutate(Detected = factor(detect_f, 
#    levels = c("Detected? Yes", "Detected? No"), 
#    labels = c("Yes", "No"))) %>% 
#ggplot(aes(x = certainty, fill = Detected)) +  
#  scale_fill_manual(values = detect_f_colors) +
#  geom_bar(position="dodge") + 
#  facet_grid(type ~ trend) +
#  theme(legend.position = "bottom") + xlab("Level of certainty")

@@

```{r}
#| label: tbl-reason
#| tbl-cap: "Proportion of participants chose each reason for their choice of plot, computed separately for each Trend and Map Type level, and whether the data plot was detected. The primary reason when participants were evaluating the choropleth map was 'consistent' or 'trend', but for the hexagon tile map it was 'clusters', when they detected the data plot."
# Qualitative analysis of reason
reasons_smry <- d %>% 
  mutate(reason = ifelse(reason =="0.0", "no reason", reason)) %>% 
  mutate(Detected = ifelse(detect_f == "Detected? Yes", 
                           "Yes", "No"),
    Trend = trend) %>% 
  group_by(Trend, Detected, type) %>% 
  count(reason) %>%
  ungroup() %>%
  group_by(Trend, Detected, type) %>%
  mutate(prop = n/sum(n)) %>% 
  ungroup() %>%
  select(Trend, Detected, type, reason, prop) %>%
  mutate(prop = round(prop, 2)) %>%
  pivot_wider(names_from = reason, 
              values_from = prop, 
              values_fill = 0) %>%
  select(Trend, Detected, type, clusters, trend, consistent, hotspots, `no reason`)

#  mutate(prop = round(n/sum(n), 2), r_prop = paste0(reason, ":", prop)) %>% 
#  top_n(1, n) %>% summarise(reasons = paste(reason, collapse=", ")) %>% 
#  pivot_wider(names_from = c("type"), values_from = c("reasons")) %>% 
#  ungroup() %>% 
reasons_smry |>  
  arrange(desc(Detected), Trend, type) |>
  knitr::kable(format = "latex", booktabs = TRUE,
               col.names = c("Trend", "Detect", "Type", 
                             "clusters", "trend", "consistent",
                             "hotspots", "none"),
                 align = "lllrrrrr",
    linesep = c("", "\\addlinespace", "", "\\addlinespace","", "\\addlinespace","", "\\addlinespace","", "\\addlinespace","")) #%>% 
  #kable_styling(bootstrap_options = "hold_position") %>% 
  #collapse_rows(., columns = 1)
```

@@

The source code to produce this document can be found on [GitHub](https://github.com/srkobakian/experiment/paper).
Supplementary materials have been included to discuss the survey procedures and the lineups that were used.
The full set of images can be found here, too.

@@

 - Document creation: quarto [@Allaire_Quarto_2025], anzjs template [@quarto-anzjs], knitr [@knitr].
 - Lineup creation: nullabor [@nullabor], gstat [@gstat].
 - Data analysis: tidyverse [@tidyverse], ggthemes [@ggthemes], RColorBrewer [@RColorBrewer].
 - Plots: ggplot2 [@ggplot2], cowplot [@cowplot], png [@png], grid [@grid], ggbeeswarm [@ggbeeswarm], ggmosaic [@ggmosaic].
 - Modelling and summary presentation: lme4 [@lme4], kableExtra [@kableExtra].

>>>>

  # Downloaded data
d <- read_xlsx("data/experiment-export.xlsx", sheet=2) |>
  filter(!is.na(contributor)) |>
  mutate(contributor = factor(contributor))

n_all_contributors <- d |> count(contributor, sort=TRUE) |> 
  summarise(n_contributors = length(contributor))

@@

# Check data set 
# Need to clean multiple entries, 48, 24
# remove duplicated entries due to submit button
d <- d |> group_by(group, contributor, image_name) |>
  slice(1) |> ungroup() |> 
  arrange(group, contributor, plot_order)

# Remove contributors who did not provide answers to most questions
keep <- d |> count(contributor, sort = TRUE) |> filter(n > 10)
d <- d |> 
  filter(contributor %in% keep$contributor) |>
  filter(contributor != "1234567890") # Testing id

# Remove contributors who did not provide any choices
# Or an insufficient amount of responses
bad_contribs <- d |> group_by(contributor) |> 
  #summarise(sum0 = sum(choice)) |> 
  summarise(n_zero = sum(if_else(choice == 0, 1, 0))) |>
  #filter(sum0 < 13) |> 
  filter(n_zero > 8) |>
  pull(contributor)

# Person is in list above
same_answer <- d |> group_by(contributor) |> 
  summarise(v = sd(choice))

@@

d <- d |> 
  filter(!(contributor %in% bad_contribs))

n_contributors <- d |> count(contributor, sort=TRUE) |> 
  summarise(n_contributors = length(contributor))

@@

d <- d |> mutate(certainty = factor(as.character(certainty),
  levels = c("1", "2", "3", "4","5"), ordered=TRUE))
```

```{r reps}
replicate <- tibble(image_name = c("aus_cities_12_geo.png", "aus_cities_12_hex.png", 
                                   "aus_cities_3_geo.png", "aus_cities_3_hex.png",
                                   "aus_cities_4_geo.png", "aus_cities_4_hex.png",
                                   "aus_cities_9_geo.png", "aus_cities_9_hex.png",
                                   "aus_nwse_2_geo.png", "aus_nwse_2_hex.png",
                                   "aus_nwse_3_geo.png", "aus_nwse_3_hex.png",
                                   "aus_nwse_5_geo.png", "aus_nwse_5_hex.png",
                                   "aus_nwse_6_geo.png", "aus_nwse_6_hex.png",
                                   "aus_three_12_geo.png", "aus_three_12_hex.png",
                                   "aus_three_5_geo.png", "aus_three_5_hex.png",
                                   "aus_three_8_geo.png", "aus_three_8_hex.png",
                                   "aus_three_9_geo.png", "aus_three_9_hex.png"),
                    replicate = c(1, 1, 2, 2, 3, 3, 4, 4, 
                                  1, 1, 2, 2, 3, 3, 4, 4,
                                  1, 1, 2, 2, 3, 3, 4, 4))
# Add rep info to data
d <- d |> left_join(replicate, by = "image_name")
```

```{r pdetection_group}
# Tidy for analysis
d <- d |> 
  separate(image_name, c("nothing", "trend", "location", "type", "extra"), remove = FALSE) |>
  select(-nothing, -extra) |>
  mutate(location = as.numeric(location), 
    # detect measures the accuracy of the choice
         detect = ifelse(location == choice, 1, 0)) |> 
  mutate(trend = case_when(
    trend == "nwse" ~ "NW-SE",
    trend == "cities" ~ "All Cities",
    trend == "three" ~ "Three Cities")) |> 
  mutate(trend = fct_relevel(trend,"All Cities","Three Cities", "NW-SE")) |> 
  mutate(type = case_when(
    type == "hex" ~"Hex.",
    TRUE~"Choro.")) |> 
    mutate(detect_f = factor(detect, levels = c(1, 0), labels = c("Detected? Yes", "Detected? No")))

plots <- d |> group_by(group, trend, type, location) |>
  # pdetect measures the aggregated accuracy of the choices
  summarise(pdetect = length(detect[detect == 1])/length(detect)) 
```

@@

sa2 <- sa22011 |> 
  filter(!st_is_empty(geometry)) |> 
  filter(!state_name_2011 == "Other Territories") |> 
  filter(!sa2_name_2011 == "Lord Howe Island")

SIR <- read_csv("data/SIR Downloadable Data.csv") |> 
  filter(SA2_name %in% sa2$sa2_name_2011) |> 
  dplyr::select(Cancer_name, SA2_name, Sex_name, p50) |> 
  filter(Cancer_name == "Thyroid", Sex_name == "Females")
ERP <- read_csv("data/ERP.csv") |>
  filter(REGIONTYPE == "SA2", Time == 2011, Region %in% SIR$SA2_name) |> 
  dplyr::select(Region, Value)
# Alternative maps
# Join with sa2 sf object
sa2thyroid_ERP <- left_join(sa2, SIR, by = c("sa2_name_2011" = "SA2_name")) |>
  left_join(ERP |> 
              dplyr::select(Region, 
              Population = Value), by = c("sa2_name_2011"= "Region")) |> 
  filter(!st_is_empty(geometry))
sa2thyroid_ERP <- sa2thyroid_ERP |> 
  #filter(!is.na(Population)) |> 
  filter(!sa2_name_2011 == "Lord Howe Island") |> 
  mutate(SIR = map_chr(p50, aus_colours)) |> 
  st_as_sf()

@@

## Check the projection uses long and lat
## Create centroids set
sa2centroids <- create_centroids(sa2, "sa2_name_2011")
## Create hexagon grid
# For speed, consider if the buffer distance beyond the centroids is necessary
grid <- create_grid(centroids = sa2centroids,
                    hex_size = 0.15,
                    buffer_dist = 4, verbose = TRUE)
## Allocate polygon centroids to hexagon grid points
# This is a time consuming process for a large data set
aus_hexmap <- allocate(
  centroids = sa2centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.15,
  hex_filter = 8,
  focal_points = capital_cities,
  width = 25,
  verbose = TRUE)
save(aus_hexmap, file = "data/aus_hexmap.rda") 
}

@@

## Prepare to plot
fort_hex <- fortify_hexagon(data = aus_hexmap,
                            sf_id = "sa2_name_2011",
                            hex_size = 0.3) |> 
            left_join(sa2thyroid_ERP |>
            select(sa2_name_2011, SIR, p50))
fort_aus <- sa2thyroid_ERP |>
  fortify_sfc()
## Make a plot
aus_hexmap_plot <- ggplot() +
  geom_polygon(aes(x = long,  y = lat,  
        group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey", 
               linewidth = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, 
                   fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal() +
  ggtitle("b. hexagon tile map")
```

@@

ggplot(expt_design, ) +
  geom_tile(aes(x=group, y=1, fill=type)) +
  geom_text(aes(x=group, y=1, label = paste("loc:", location))) + 
  facet_grid(replicate~trend, axes = "all", labeller = labeller(replicate = label_both, trend = label_both)) +
  ylab("") +
  scale_fill_manual("", values = type_colors) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks = element_blank())
```

@@

var.sim1 <- predict(var.g.dummy, newdata = sa3_centroids, nsim = 12) |> 
  left_join(sa3_centroids, ., by=c("longitude", "latitude"))

sa3_sims1_1 <- as_tibble(var.sim1) |> 
  mutate_at(sims, ~map_dbl(1:nrow(sa3), spatial_smoother, 
                           values_vector = ., area_weight = 0.5, neighbours_list = sa3_neighbours))

@@

```{r}
#| label: fig-detect-compare
#| fig-cap: "The detection rates achieved by participants are contrasted when viewing the four replicates of the three trend models. Each point shows the probability of detection for the lineup display, the facets separate the trend models hidden in the lineup. The points for the same data set shown in a choroleth or hexagon tile map display are linked to show the difference in the detection rate."
#| fig-height: 4
#| fig-width: 6
#| out-width: 100%
# Detectability rate for each lineup (image)
d_smry <- d |> group_by(trend, type, replicate) |>
  # pdetect measures the aggregated accuracy of the choices
  summarise(pdetect = length(detect[detect == 1])/length(detect)) |>
  ungroup()

# Numerical summary
diffs <- d_smry |> spread(type, pdetect) |>
  mutate(dif = `Hex.` - `Choro.`)

@@

```{r}
#| eval: false
# Check detection rate across groups
d |> group_by(group) |>
    # pdetect measures the aggregated accuracy of the choices
    summarise(pdetect = length(detect[detect == 1])/length(detect)) |>
    ungroup()
```

@@

d |> group_by(type, trend) |>
  summarise(m = as.character(round(mean(detect), 2)),
      std.dev = as.character(round(sd(detect), 2))) |> 
  mutate(std.dev = ifelse(std.dev == 0.5, "(0.50)", paste0("(", std.dev, ")"))) |>   mutate(m = ifelse(m == 0.4, "0.40", m)) |> 
  gather(stat, value, m, std.dev) |> 
  pivot_wider(names_from = "trend", values_from = "value") |> 
  arrange(type) |> ungroup() |> 
  mutate(Type = types) |> select(Type, `NW-SE`, `Three Cities`, `All Cities`) |>
  knitr::kable(., format = "latex", align = "lccc", booktabs = TRUE, 
    linesep = c("", "\\addlinespace"))
```

@@

detection_rates <- tidy(detect_fit) |>
  mutate(detection_rates = round(exp(estimate)/(1+exp(estimate)),2)) |>
  select(term, estimate, detection_rates) |> pull(detection_rates)

@@

tidy(detect_fit) |>
  mutate_at(.vars = c("estimate", "std.error"), round, 2) |> 
  mutate(p.value = round(p.value, digits=2)) |> 
  rowwise() |>
  mutate(sig = case_when(
  p.value <= 0.001 ~ "$^{***}$",
  p.value <= 0.01 ~  "$^{**}$",
  p.value <= 0.05 ~  "$^{*}$",
  p.value <= 0.01 ~  "$^{.}$",
  TRUE ~ "$^{ }$")) |>
  ungroup() |> 
  filter(!is.na(std.error)) |>
  mutate(term = glmer_terms) |> 
  select(Term = term, 
    Est. = estimate,
    `Std. Err.` = std.error, 
    `P-val.` = p.value, 
    Sig. = sig) |> 
  knitr::kable(format = "latex", escape = FALSE, align= "rrlrr", 
    booktabs = T, linesep = c("", "\\addlinespace", "", "\\addlinespace", ""))
```

@@

```{r}
#| label: fig-beeswarm
#| fig-cap: "The distribution of the time taken (seconds) to submit a response for each combination of trend, whether the data plot was detected, and type of display, shown using horizontally jittered dotplots. The coloured point indicates average time taken for each map type. Although some participants take just a few seconds per evaluation, and some take as much as 60 seconds, but there is very little difference in time taken between map types."
#| fig-height: 4
#| fig-width: 6
#| out-width: 100%
s <- d |> group_by(type, trend, detect_f) |>
  summarise(m=median(time_taken), 
            q1=quantile(time_taken, 0.25), 
            q3=quantile(time_taken, 0.75)) 
ggplot() + 
  geom_quasirandom(data=d, aes(x=type, y=time_taken), alpha=0.9) + 
  #geom_hline(data=s, aes(yintercept=m, color=type)) +
  geom_point(data=s, aes(x=type, y=m, color=type), size=3, alpha=0.7) +
  #geom_errorbar(data=s, aes(x=type, ymin=q1, ymax=q3, color=type), width=0.3, size=2) +
  scale_color_manual("", values = type_colors) +
  facet_grid(detect_f~trend) + 
  ylab("Time taken (seconds)") + xlab("") +
  theme(legend.position="none")
```

```{r}
#| eval: false
# Check groups and demographics
d |> group_by(group) |>   
  summarise(m=mean(time_taken), 
            s=sd(time_taken))
d |> group_by(gender) |>   
  summarise(m=mean(time_taken), 
            s=sd(time_taken))
d |> group_by(age) |>   
  summarise(m=mean(time_taken), 
            s=sd(time_taken))
d |> group_by(education) |>   
  summarise(m=mean(time_taken), 
            s=sd(time_taken))

@@

```{r}
# Check time difference between default choice
d_zeros <- d |> mutate(zeros = if_else(choice == 0, 0, 1)) |>
  group_by(zeros) |>   
  summarise(m=median(time_taken), 
            mn=min(time_taken),
            mx=max(time_taken),
            q1=quantile(time_taken, 0.25), 
            q3=quantile(time_taken, 0.75),
            n=n())

d_zeros_detect <- d |> mutate(zeros = if_else(choice == 0, 0, 1)) |>
  group_by(zeros, detect_f) |>   
  summarise(m=median(time_taken), 
            mn=min(time_taken),
            mx=max(time_taken),
            q1=quantile(time_taken, 0.25), 
            q3=quantile(time_taken, 0.75),
            n=n())
# Summarise across subjects
contributor_smry <- d |> group_by(contributor) |>   
  summarise(m_d=mean(detect), 
            s_d=sd(detect),
            m_t=mean(time_taken), 
            s_t=sd(time_taken)) 
```

@@

```{r}
#| label: fig-certainty 
#| fig-cap: "The distribution of certainty chosen by participants when viewing hexagon tile map or choropleth displays, shown as a mosaic plot, faceted by the trend model and whether the plot was detected or not. At most trend models and whether the plot was detected or not, participants were more likely to choose a high certainty when evaluating a choropleth map."
#| fig-height: 5
#| fig-width: 10
#| out-width: 100%
d <- d |> 
  mutate(certainty = as_factor(certainty)) |> 
  mutate(replicate_f = as_factor(replicate))

#d |> 
#  mutate(Detected = factor(detect_f, 
#    levels = c("Detected? Yes", "Detected? No"), 
#    labels = c("Yes", "No"))) |> 
#ggplot(aes(x = certainty, fill = Detected)) +  
#  scale_fill_manual(values = detect_f_colors) +
#  geom_bar(position="dodge") + 
#  facet_grid(type ~ trend) +
#  theme(legend.position = "bottom") + xlab("Level of certainty")

@@

```{r}
#| label: tbl-reason
#| tbl-cap: "Proportion of participants chose each reason for their choice of plot, computed separately for each Trend and Map Type level, and whether the data plot was detected. The primary reason when participants were evaluating the choropleth map was 'consistent' or 'trend', but for the hexagon tile map it was 'clusters', when they detected the data plot."
# Qualitative analysis of reason
reasons_smry <- d |> 
  mutate(reason = ifelse(reason =="0.0", "no reason", reason)) |> 
  mutate(Detected = ifelse(detect_f == "Detected? Yes", 
                           "Yes", "No"),
    Trend = trend) |> 
  group_by(Trend, Detected, type) |> 
  count(reason) |>
  ungroup() |>
  group_by(Trend, Detected, type) |>
  mutate(prop = n/sum(n)) |> 
  ungroup() |>
  select(Trend, Detected, type, reason, prop) |>
  mutate(prop = round(prop, 2)) |>
  pivot_wider(names_from = reason, 
              values_from = prop, 
              values_fill = 0) |>
  select(Trend, Detected, type, clusters, trend, consistent, hotspots, `no reason`)

#  mutate(prop = round(n/sum(n), 2), r_prop = paste0(reason, ":", prop)) |> 
#  top_n(1, n) |> summarise(reasons = paste(reason, collapse=", ")) |> 
#  pivot_wider(names_from = c("type"), values_from = c("reasons")) |> 
#  ungroup() |> 
reasons_smry |>  
  arrange(desc(Detected), Trend, type) |>
  knitr::kable(format = "latex", booktabs = TRUE,
               col.names = c("Trend", "Detect", "Type", 
                             "clusters", "trend", "consistent",
                             "hotspots", "none"),
                 align = "lllrrrrr",
    linesep = c("", "\\addlinespace", "", "\\addlinespace","", "\\addlinespace","", "\\addlinespace","", "\\addlinespace","")) #|> 
  #kable_styling(bootstrap_options = "hold_position") |> 
  #collapse_rows(., columns = 1)
```

@@

This document was written using quarto. This document contains the code to produce the summaries, plots and additional checks in the paper.  and all the code to reproduce the analysis, and do additional checks can be found on [GitHub](https://github.com/srkobakian/experiment/paper).
Supplementary materials have been included to discuss the survey procedures and the lineups that were used.
The full set of images can be found here, too.

@@

 - Document creation: quarto [@Allaire_Quarto_2025], anzjs template [@quarto-anzjs], knitr [@knitr].
 - Lineup creation: nullabor [@nullabor], gstat [@gstat].
 - Data analysis: tidyverse [@tidyverse], ggthemes [@ggthemes], RColorBrewer [@RColorBrewer].
 - Plots: ggplot2 [@ggplot2], cowplot [@cowplot], png [@png], ggbeeswarm [@ggbeeswarm], ggmosaic [@ggmosaic].
 - Modelling and summary presentation: lme4 [@lme4], kableExtra [@kableExtra].
